Table-Driven Customization of Web Applications

### 2020-05-19 paper.tex
setup pipeline to google docs and pdf

TBD, Summarize the paper

@@

\begin{itemize}
\tightlist
\item
  use airbnb + expedia
\item
  instacart?
\item
  new screenshots
\item
  mention in text all the different sites we've used it with
\item
  1 externally contributed adapter
\item
  flux story
\end{itemize}

@@

\hypertarget{specific-table-adapters}{%
\subsection{Specific table adapters}\label{specific-table-adapters}}

\emph{(Todo: maybe move this upwards into the worked example?)}

Currently, our system supports three types of table adapters.

\textbf{DOM scraping adapter}: extracts data from the DOM of a web page,
and manipulates the DOM to re-order rows, edit form entries, and inject
annotations. A programmer only needs to write a single function which
returns scraped data and pointers to relevant DOM elements; the Wildcard
framework uses this function to implement the rest of the needed
functionality.

\textbf{AJAX scraping adapter}: intercepts AJAX requests made by a web
page, and extracts information from those requests. A programmer writes
a function which specifies how to extract data from an AJAX request, and
the framework handles the details of actually intercepting requests and
calling the programmer-defined function.

\textbf{Local storage adapter}: simply stores a table of data in the
browser.

@@

\hypertarget{table-editor-instruments}{%
\subsection{Table editor instruments}\label{table-editor-instruments}}

Our editor is closer to visual SQL query engines like Sieuferd or
Airtable than a freeform spreadsheet.

\hypertarget{future-table-adapters}{%
\subsection{Future table adapters}\label{future-table-adapters}}

However, we have intentionally designed the table adapter API to be
general enough to encapsulate other types of data and additional
functionality in the future. Here are three concrete examples of such
possibilities:

@@

With data-oriented customization and Wildcard, we use a hybrid approach
that takes the best of both worlds. Programmers implement an API wrapper
that is internally implemented as a black-box customization, but
externally provides a semantic interface to the application. End users
get an ergonomic and simplified customization experience, but without
the need to depend on a first-party application developer exposing
extension APIs.

@@

borrow from the in-place

@@

>>>>

Many software applications don't meet the precise needs of their users.
When this happens, the only recourse is to complain to the developers,
or, more likely, to simply give up. Back in 1977, in Personal Dynamic
Media \citep{kay1977}, Alan Kay envisioned personal computing as a
medium that let a user ``mold and channel its power to his own needs,''
but today, software behaves more like concrete than clay.

\begin{itemize}
\tightlist
\item
  many approaches to customizing software. Most are imperative command
  based. Can't get around it.
\item
  there are successful systems for visual SQL queries: Airtable,
  Sieuferd. Even spreadsheets.
\item
  what if you could see the data queries powering a website and directly
  modify them yourself?
\item
  we call this table-driven customization.
\end{itemize}

To make it work with the real world, we develop Wildcard, shim this on
top of existing web apps. Show in Section Examples that it really works!
can do tons of useful stuff with real apps.

Architecture: Describe implementation of Wildcard. We've developed
generic abstractions that make this system even more extensible in the
future. we also propose a future architecture where apps are built for
this; they expose their data queries and their structured data results.

Design Principles: explain ideas behind the system, generalizable to
software customization more generally

Related Work: table editors,

@@

\emph{This section still todo. Borrow heavily from Convivial paper, but
extend with newer results}

\begin{itemize}
\tightlist
\item
  Just use airbnb + expedia from Convivial paper?
\item
  instacart + hacker news?
\item
  summarize/mention all 12 sites we've used it with
\item
  real usage:

  \begin{itemize}
  \tightlist
  \item
    1 externally contributed adapter
  \item
    flux adapter in actual use by Sergio
  \end{itemize}
\end{itemize}

@@

\hypertarget{table-adapters}{%
\subsection{Table adapters}\label{table-adapters}}

Here we describe the three types of table adapters which we have
implemented so far to power the customizations shown above. Then, to
demonstrate the generality of the table adapter paradigm, we describe
three more hypothetical types of table adapters which could extend the
power of the Wildcard system.

\hypertarget{existing-adapters}{%
\subsubsection{Existing Adapters}\label{existing-adapters}}

A \textbf{DOM scraping adapter} extracts data from the DOM of a web
page, and manipulates the DOM to re-order rows, edit form entries, and
inject annotations. A programmer only needs to write a single function
which returns scraped data and pointers to relevant DOM elements; the
Wildcard framework uses this function to implement the rest of the
needed functionality.

(\emph{do we want to call DOM scraping adapter a ``live'' adapter
because its data is also shown elsewhere\ldots?})

An \textbf{AJAX scraping adapter} intercepts AJAX requests made by a web
page, and extracts information from those requests. A programmer writes
a function which specifies how to extract data from an AJAX request, and
the framework handles the details of actually intercepting requests and
calling the programmer-defined function.

The \textbf{local storage adapter} simply stores a table of data in the
browser.

\hypertarget{future-adapters}{%
\subsubsection{Future Adapters}\label{future-adapters}}

we have intentionally designed the table adapter API to be general
enough to encapsulate other types of data and additional functionality
in the future. Here are three concrete examples of such possibilities:

@@

\begin{itemize}
\tightlist
\item
  formulas: do they go here?
\end{itemize}

\hypertarget{table-editor}{%
\subsection{Table editor}\label{table-editor}}

\begin{itemize}
\tightlist
\item
  Our editor is closer to visual SQL query engines like Sieuferd or
  Airtable than a freeform spreadsheet.
\item
  cell editors
\item
  Note that there could be other entire table editors
\end{itemize}

@@

With Wildcard, we use a hybrid approach that takes the best of both
worlds. Programmers implement an API wrapper that is internally
implemented as a black-box customization, but externally provides a
semantic interface to the application. End users get an ergonomic and
simplified customization experience, but without the need to depend on a
first-party application developer exposing extension APIs.

@@

\emph{borrow from convivial paper's discussion of in-place}

@@

\hypertarget{related-work}{%
\section{Related Work}\label{related-work}}

Find a way to organize all this:

\begin{itemize}
\tightlist
\item
  our own workshop paper. since then\ldots{}

  \begin{itemize}
  \tightlist
  \item
    fundamentally rearchitected around table adapters
  \item
    evaluated on many more websites
  \item
    more fully describing how the system works
  \end{itemize}
\item
  browser extensions
\item
  instrumental interaction, Scotty, Webstrates
\item
  customization research tools: Chickenfoot, Coscripter

  \begin{itemize}
  \tightlist
  \item
    Wildcard, like Chickenfoot, wants to hide HTML from users. But we
    show a structured data view, whereas Chickenfoot shows nothing
  \end{itemize}
\item
  desktop customization: Applescript, VBA, COM
\item
  browser dev tools
\item
  database GUIs
\item
  spreadsheets
\end{itemize}
